### 2020-01-28 wuhan.tex
stuff

\section*{Early estimates of $\mathcal R_0$}

Early in an outbreak, $\mathcal R_0$ cannot be estimated directly.
Instead, estimates of $\mathcal R_0$ are often inferred from
the exponential growth rate $r$ can be estimated reliably from incidence data.
Given estimates of the exponential growth rate $r$ and the distribution $g(\tau)$ of 
generation intervals (i.e., the time between when a person become 
infected and that person infects another person), the basic reproduction
number can be estimated via the Euler-Lotka equation:
\begin{equation}
1/\mathcal R_0 = \int \exp(-r\tau) g(\tau) d\tau.
\end{equation}
In other words, estimates of $\mathcal R_0$
necessarily depends on the assumptions about 
growth rate $r$ and the shape of generation-interval distribution $g(\tau)$.

@@

We use the gamma approximation framework to demonstrate
how assumptions about $r$ and $g(\tau)$
affects $\mathcal R_0$.
Assuming that generation intervals follow a gamma distribution 
with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
we have
\begin{equation}
\mathcal R_0 = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
\end{equation}
This equation demonstrates that generation-interval distributions
with a larger mean (higher $\bar{G}$) or less variability (lower $\kappa$)
will give a higher estimate of $\mathcal R_0$.
Likewise, strong assumptions about $\bar{G}$ and $\kappa$ will give
estimates of $\mathcal R_0$ with narrow confidence intervals.

Figure 1 clearly demonstrates the lack of uncertainties in the underlying parameters
of the current $\mathcal R_0$ estimates.
In particular, no studies properly propagate uncertainties associated with the
amount of variability in generation intervals, which can have 
large effects on the estimates of $\mathcal R_0$
For example, when the relative mean generation interval is long ($r\bar{G}\approx 2$ based on the estimates by \cite{zhaoncov}),
varying $\kappa$ from 0 to 1 reduces the estimate of $\mathcal R_0$ by more than a two fold.

\begin{equation}
\begin{aligned}
\mathcal R_i &\sim \textrm{distrib}(\theta_i)\\
\bar G_i &\sim \textrm{distrib}(\theta_i)\\
\kappa &\sim \textrm{distrib}(\theta_i)\\
\end{aligned}
\end{equation}

>>>>

\section{Methods and results}

Early in an outbreak, $\mathcal R_0$ cannot be estimated directly.
Instead, estimates of $\mathcal R_0$ are often inferred from
the exponential growth rate $r$, which can be estimated reliably from incidence data.
Given estimates of the exponential growth rate $r$ and the distribution $g(\tau)$ of
generation intervals (i.e., the time between when a person become 
infected and that person infects another person), the basic reproduction
number can be estimated via the Euler-Lotka equation:
\begin{equation}
1/\mathcal R_0 = \int \exp(-r\tau) g(\tau) d\tau.
\end{equation}
In other words, estimates of $\mathcal R_0$
will depend on the assumptions about the
exponential growth rate $r$ and the shape of generation-interval distribution $g(\tau)$.

@@

We use the gamma approximation framework to first characterize the
amount of uncertainty present in the underlying parameters and then 
to assess the 
degree to which these uncertainties
affect the estimate of $\mathcal R_0$.
Assuming that generation intervals follow a gamma distribution 
with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
we have
\begin{equation}
\mathcal R_0 = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
\end{equation}
This equation demonstrates that a generation-interval distribution
that assumes a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$)
will give a higher estimate of $\mathcal R_0$ for the same value of $r$.

First, we gathered information on estimates of $\mathcal R_0$ and their
assumptions about the underlying generation-interval distributions from preprint servers (e.g., bioRxiv, medRxiv, and SSRN).
Then, we model each parameter ($\mathcal R_0$, $\bar G$, and $\kappa$) with an appropriate probability distribution (either uniform or gamma) that matches the reported estimates or assumptions.
For example, \cite{imaincov} estimated $\mathcal R_0 = 2.5$
